---
title: "Reading zipped data"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(zip)
```

# Call the Physician and Other Supplier

Sources can be found from here: https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Medicare-Provider-Charge-Data/Physician-and-Other-Supplier

This has data for the years

- 2017
- 2016
- 2015
- 2014
- 2013
- 2012

Note that when you click the download link, in the browser you need to extract the exact URL. E.g. on [this page](https://www.cms.gov/apps/ama/license.asp?file=http://download.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Medicare-Provider-Charge-Data/Downloads/Medicare_Provider_Util_Payment_PUF_CY2017.zip), look in the URL to find the address of the actual file:

> http://download.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Medicare-Provider-Charge-Data/Downloads/Medicare_Provider_Util_Payment_PUF_CY2017.zip

## For year 2017

```{r get2017, eval=F}
url <- "http://download.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Medicare-Provider-Charge-Data/Downloads/Medicare_Provider_Util_Payment_PUF_CY2017.zip"

file_name <- "Medicare_Provider_Util_Payment_PUF_CY2017.txt"




temp <- tempfile()
download.file(url, temp) # download to temp
file_uz <- unz(temp, file_name)
data <- readr::read_delim(file_uz, "\t", escape_double=F, trim_ws=T)
unlink(temp)
  
# zip::unzip(temp, list=T)
rm(url, file_name, file_uz, temp)

data %>% write_csv("data/PUP_2017.csv")
```

```{r writeMerged17}
data <- read_csv("data/PUP_2017.csv")
# Interventional Radiology
# Neurosurgery
# Vascular Surgery

IR <- data %>%
  filter(provider_type=="Interventional Radiology") %>%
  group_by(hcpcs_code, hcpcs_description, npi) %>%
  summarise(line_srvc_cnt = sum(line_srvc_cnt)) %>%  # sum of hcpcs_code & provider
  
  summarise(line_srvc_cnt = sum(line_srvc_cnt),
            NPIs          = n())

NSG <- data %>%
  filter(provider_type=="Neurosurgery") %>%
  group_by(hcpcs_code, hcpcs_description, npi) %>%
  summarise(line_srvc_cnt = sum(line_srvc_cnt)) %>%  # sum of hcpcs_code & provider
  
  summarise(line_srvc_cnt = sum(line_srvc_cnt),
            NPIs          = n())


Vasc <- data %>%
  filter(provider_type=="Vascular Surgery") %>%
  group_by(hcpcs_code, hcpcs_description, npi) %>%
  summarise(line_srvc_cnt = sum(line_srvc_cnt)) %>%  # sum of hcpcs_code & provider
  
  summarise(line_srvc_cnt = sum(line_srvc_cnt),
            NPIs          = n())
  
merged <- bind_rows(
  mutate(IR,   ProvType = "IR"),
  mutate(NSG,  ProvType = "NSG"),
  mutate(Vasc, ProvType = "Vasc")
) %>%
  mutate(Year = 2017)

merged %>% write_csv("~/Neuro_IR/data/merged17.csv")
rm(data)
```

```{r}
merged %>%
  filter(hcpcs_code %in% IR$hcpcs_code,
         hcpcs_code %in% NSG$hcpcs_code,
         ProvType != "Vasc") %>%
  select(-NPIs) %>%
  spread(ProvType, line_srvc_cnt) %>%
  mutate(PctNSG = NSG/(NSG+IR)) %>%
  arrange(PctNSG)
  # arrange(hcpcs_code, ProvType) %>%
  # View()
```


## For year 2016

```{r get2016, eval=F}
url <- "https://downloads.cms.gov/files/Medicare-Physician-and-Other-Supplier-PUF.zip"
temp <- tempfile()
download.file(url, temp) # download to temp

# Find the name of the file
zip::zip_list(temp)
file_name <- "Medicare_Provider_Util_Payment_PUF_CY2016.txt"


# Load as dataset
file_uz <- unz(temp, file_name)
data <- readr::read_delim(file_uz, "\t", escape_double=F, trim_ws=T)


# write to csv
data %>% 
  filter(NPI!="0000000001") %>%
  write_csv("data/PUP_2016.csv")


# clean up
unlink(temp)
rm(url, file_name, file_uz, temp)

```

```{r writeMerged16}
data <- read_csv("data/PUP_2016.csv")

IR <- data %>%
  filter(PROVIDER_TYPE=="Interventional Radiology") %>%
  group_by(HCPCS_CODE, HCPCS_DESCRIPTION, NPI) %>%
  summarise(LINE_SRVC_CNT = sum(LINE_SRVC_CNT)) %>%  # sum of hcpcs_code & provider
  
  summarise(LINE_SRVC_CNT = sum(LINE_SRVC_CNT),
            NPIs          = n())

NSG <- data %>%
  filter(PROVIDER_TYPE=="Neurosurgery") %>%
  group_by(HCPCS_CODE, HCPCS_DESCRIPTION, NPI) %>%
  summarise(LINE_SRVC_CNT = sum(LINE_SRVC_CNT)) %>%  # sum of hcpcs_code & provider
  
  summarise(LINE_SRVC_CNT = sum(LINE_SRVC_CNT),
            NPIs          = n())


Vasc <- data %>%
  filter(PROVIDER_TYPE=="Vascular Surgery") %>%
  group_by(HCPCS_CODE, HCPCS_DESCRIPTION, NPI) %>%
  summarise(LINE_SRVC_CNT = sum(LINE_SRVC_CNT)) %>%  # sum of hcpcs_code & provider
  
  summarise(LINE_SRVC_CNT = sum(LINE_SRVC_CNT),
            NPIs          = n())
  
merged <- bind_rows(
  mutate(IR,   ProvType = "IR"),
  mutate(NSG,  ProvType = "NSG"),
  mutate(Vasc, ProvType = "Vasc")
) %>%
  mutate(Year = 2016)

merged %>% write_csv("~/Neuro_IR/data/merged16.csv")
rm(data)
```

## For year 2015

```{r get2015, eval=F}
url <- "http://download.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Medicare-Provider-Charge-Data/Downloads/Medicare_Provider_Util_Payment_PUF_CY2015.zip"
temp <- tempfile()
download.file(url, temp) # download to temp

# Find the name of the file
zip::zip_list(temp)
file_name <- "Medicare_Provider_Util_Payment_PUF_CY2015.txt"


# Load as dataset
file_uz <- unz(temp, file_name)
data <- readr::read_delim(file_uz, "\t", escape_double=F, trim_ws=T)


# write to csv
data %>% 
  filter(npi!="0000000001") %>%
  write_csv("data/PUP_2015.csv")


# clean up
unlink(temp)
rm(url, file_name, file_uz, temp)

```

```{r writeMerged15, eval=F}
data <- read_csv("data/PUP_2015.csv")

IR <- data %>%
  filter(provider_type=="Interventional Radiology") %>%
  group_by(hcpcs_code, hcpcs_description, npi) %>%
  summarise(line_srvc_cnt = sum(line_srvc_cnt)) %>%  # sum of hcpcs_code & provider
  
  summarise(line_srvc_cnt = sum(line_srvc_cnt),
            NPIs          = n())

NSG <- data %>%
  filter(provider_type=="Neurosurgery") %>%
  group_by(hcpcs_code, hcpcs_description, npi) %>%
  summarise(line_srvc_cnt = sum(line_srvc_cnt)) %>%  # sum of hcpcs_code & provider
  
  summarise(line_srvc_cnt = sum(line_srvc_cnt),
            NPIs          = n())


Vasc <- data %>%
  filter(provider_type=="Vascular Surgery") %>%
  group_by(hcpcs_code, hcpcs_description, npi) %>%
  summarise(line_srvc_cnt = sum(line_srvc_cnt)) %>%  # sum of hcpcs_code & provider
  
  summarise(line_srvc_cnt = sum(line_srvc_cnt),
            NPIs          = n())
  
merged <- bind_rows(
  mutate(IR,   ProvType = "IR"),
  mutate(NSG,  ProvType = "NSG"),
  mutate(Vasc, ProvType = "Vasc")
) %>%
  mutate(Year = 2015)

merged %>% write_csv("~/Neuro_IR/data/merged15.csv")
rm(data)
```


# Full NPI dataset

Data source: http://download.cms.gov/nppes/NPI_Files.html

Documentation can be found on this server under the `Full NPI documentation` folder